brados: declarative, programmable object storage

### 2016-05-25 paper.tex

To demonstrate the use of the \emph{brados} language we will show how a
high-performance shared-log based on the CORFU protocol can be built in Ceph, as well
as existing real-world interfaces used in the Ceph storage system. First we
discuss the Ceph storage system which our prototype is built upon and the 
motivating system, CORFU. We then present the challenges that programmers face
when navigating the design space during the process of building an
application-storage interface. The \emph{brados} language is then described
using the CORFU system and an existing Ceph-specific interface as motivating examples.
We conclude by discussing the optimization opportunities that we can realize
using analysis of interfaces. We hope you enjoy the show.

@@

\paragraph*{CORFU and Storage Programmability}
The second property of CORFU relevant in the context of Ceph is the dependency
CORFU places on custom storage device interfaces used to guarantee
serialization during failure and reconfiguration. Each flash device in a CORFU
cluster exposes a 64-bit write-once address space consisting of the primary I/O
interfaces \emph{write(pos, data)} and \emph{read(pos)} for accessing log
entries, as well as \emph{fill(pos)} and \emph{trim(pos)} that invalidate and
reclaim log entries, respectively. All I/O operations in CORFU initiated by
clients are tagged with an \emph{epoch} value, and flash devices are expected
to reject client requests that contain an old epoch value. To facilitate
recovery or handle system reconfiguration in CORFU, the storage devices are
also required to support a \emph{seal(epoch)} command that stores the latest
epoch and returns the maximum position written to that device. The seal
interface is used following the failure of a sequencer to calculate the tail of
the log that the sequencer should use to repopulate its in-memory state.

While the authors of the CORFU paper describe prototype device interfaces
implemented as both host-based and FPGA-based solutions, RADOS \emph{directly} supports
the creation of logical storage devices through its object class feature
described previously in Section~\ref{sec:objclass}. Thus, by using
software-based object interfaces offered by RADOS flash devices in CORFU can be
replaced by software-defined storage offering significant flexibility and a
simplified design.

@@

As we have seen, Ceph provides a rich storage API and places few restrictions
on the structure of applications. Thus a primary concern when implementing an
object interface in Ceph is deciding how native interfaces are composed into
compound operations in order to implement the semantics of the target
interface. These types of decisions are commonly referred to as
physical design, and can affect performance and application flexibility.
As we will see the design space that developers must operate in is often
large, and its dynamic nature can lead to design decisions that become
obsolete and non-optimal.

@@

  \node[initial left,state] (A)              {$EG$};
  \node[state]         (B) [below right of=A] {$CP$};
  \node[state]         (D) [right of=B]       {$IO$};
  \node[state]         (C) [above right of=A] {$GC$};
  \node[state]         (E) [right of=C]       {$US$};
  \node[state]         (F) [below right of=E] {$Fi$};

@@

We begin the process of exploring the design space by exploiting the fact that
metadata management is a complexity and performance overhead that a full
implementation must incur beyond the costs of storing log entry data.
Therefore we first explore the design space by restricting the
space to log entry addressing and log entry storage (the \emph{I/O} action
shown in Figure~\ref{fig:corfu-sm}).  These two dimensions are
represented by the first two columns of Table~\ref{tab:pd-map} which describes
the entire design space. In the I/O column \emph{KV} corresponds to the
key-value interface, and the bytestream interface is represented by \emph{AP}
for an append strategy, and \emph{EX} for a strategy that writes to an
explicit bytestream offset (both described shortly).

@@

The top row of Figure~\ref{fig:jewel_v_firefly_v_es} shows the expected
performance of different sized log appends without metadata management overhead
using each of the five strategies defined in Table~\ref{tab:pd-map}.  Starting
with Jewel and 4KB entry sizes, we see that both one-to-one strategies have
relatively poor performance. Jewel with 1KB entry sizes sees a 2.5\(\times\)
performance increase for the N:1 KV mapping. The large period of reduced
throughput for corresponds to the OSD splitting file system directories in
order to maintain a maximum directory size, and will occur in \emph{1:1, KV}
although the threshold number of objects is not reached in this example due to
the reduced overall throughput of the \emph{1:1, KV} strategy.

@@

\begin{figure}[t]
      \centering
      \includegraphics[width=0.85\linewidth]{experiments/basic-cls-rand-read/output.read.60min.1k.png}
      \caption{The expected random read performance from a log with 1KB entries
      performs best on the same strategy that the write performance performs
      best on - the N:1 bystream append strategy.}
      \label{fig:vanilla_rd_jewel}
\end{figure}

@@

Ceph is designed to run on a wide variety of commodity hardware as well as new
NVMe devices. All these devices have their own set of characteristics and
tunables (e.g., the IO operation scheduler type). In our experiments, we tested
SSD, HDDs, NVMe devices and discovered a wide range of behaviors and
performance profiles. As an example, Figure~\ref{fig:jewel-hdd-128b} shows the
write performance of 128 byte log entries using Jewel and a single HDD.
Performance is 10\(\times\) slower than its SSD counterpart in
Figure~\ref{ig:jewel_v_firefly_v_es} (top row, third column) but the behavior
and relative performance make this hardware configuration especially tricky.

The behavior of the 1:1 implementations shows throughput drops lasting for
minutes at a time -- this limits our focus to the N:1 implementations. The
performance of the N:1 bystestream implementation is almost identical to the
N:1 KV implementation (within 1\% mean throughput). Regardless of the true
bottleneck, it is clear that choosing the N:1 KV implementation is the better
choice because of the resulting implementation should be less complex and there
is minimal performance degradation.

>>>>

To demonstrate the use of the \emph{brados} language we will show how a
high-performance shared-log based on the CORFU protocol can be built in Ceph, as well
as existing real-world interfaces used in the Ceph storage system. First we
discuss the Ceph storage system which our prototype is built upon and the 
motivating system, CORFU. We then present the challenges that programmers face
when navigating the design space during the process of building an
application-storage interface. The \emph{brados} language is then described
using the CORFU system and an existing Ceph-specific interface as motivating examples.
We conclude by discussing the optimization opportunities that we can realize
using analysis of interfaces.

@@

The second property of CORFU relevant in the context of Ceph is the dependency
CORFU places on custom storage device interfaces used to guarantee
serialization during failure and reconfiguration. Each flash device in a CORFU
cluster exposes a 64-bit write-once address space consisting of the primary I/O
interfaces \emph{write(pos, data)} and \emph{read(pos)} for accessing log
entries, as well as \emph{fill(pos)} and \emph{trim(pos)} that invalidate and
reclaim log entries, respectively. All I/O operations in CORFU initiated by
clients are tagged with an \emph{epoch} value, and flash devices are expected
to reject client requests that contain an old epoch value. To facilitate
recovery or handle system reconfiguration in CORFU, the storage devices are
also required to support a \emph{seal(epoch)} command that stores the latest
epoch and returns the maximum position written to that device. The seal
interface is used following the failure of a sequencer to calculate the tail of
the log that the sequencer should use to repopulate its in-memory state.

\paragraph*{Storage Programmability}
While the authors of the CORFU paper describe prototype device interfaces
implemented as both host-based and FPGA-based solutions, RADOS \emph{directly} supports
the creation of logical storage devices through its object class feature
described in Section~\ref{sec:objclass}. Thus, by using
software-based object interfaces offered by RADOS flash devices in CORFU can be
replaced by software-defined interfaces offering significant flexibility and a
simplified design.

@@

As we have seen, Ceph provides a rich storage API and places few requirements
on the structure of applications. Thus a primary concern when implementing an
object interface in Ceph is deciding how native interfaces are composed into
compound operations in order to implement the semantics of the target
interface. These types of decisions are commonly referred to as
physical design, and can affect performance and application flexibility.
As we will see the design space that developers must operate in is often
large, and its dynamic nature can lead to design decisions that become
obsolete and non-optimal.

@@

  \node[initial left,state] (A)              {$EG$};
  \node[state]         (B) [below right of=A] {$CP$};
  \node[state]         (D) [right of=B]       {$IO$};
  \node[state]         (C) [above right of=A] {$GC$};
  \node[state]         (E) [right of=C]       {$US$};
  \node[state]         (F) [below right of=E] {$Out$};

@@

We begin the process of exploring the design space by focusing on log entry
addressing and log entry storage (the \emph{I/O} action shown in
Figure~\ref{fig:corfu-sm}).  These two dimensions are represented by the first
two columns of Table~\ref{tab:pd-map} which describes the entire design space.
In the I/O column \emph{KV} corresponds to the key-value interface, and the
bytestream interface is represented by \emph{AP} for an append strategy, and
\emph{EX} for a strategy that writes to an explicit bytestream offset (both
described shortly).

@@

The top row of Figure~\ref{fig:jewel_v_firefly_v_es} shows the expected
performance of different sized log appends without metadata management overhead
using each of the five strategies defined in Table~\ref{tab:pd-map}.  Starting
with Jewel and 4KB entry sizes, we see that both one-to-one strategies have
relatively poor performance. Jewel with 1KB entry sizes sees a 2.5\(\times\)
performance increase for the \texttt{(N:1,KV)} mapping. The large period of reduced
throughput for corresponds to the OSD splitting file system directories in
order to maintain a maximum directory size, and will occur in \texttt{(1:1, KV)}
although the threshold number of objects is not reached in this example due to
the reduced overall throughput of the \texttt{(1:1, KV)} strategy.

@@

\begin{figure}[t]
      \centering
      \includegraphics[width=0.85\linewidth]{experiments/basic-cls-rand-read/output.read.60min.1k.png}
      \caption{The expected random read performance from a log with 1KB entries
      performs best on the same strategy that the writes perform
      best on - the \texttt{(N:1, BS)} append strategy.}
      \label{fig:vanilla_rd_jewel}
\end{figure}

@@

Ceph is designed to run on a wide variety of commodity hardware as well as new
NVMe devices. All these devices have their own set of characteristics and
tunables (e.g., the IO operation scheduler type). In our experiments, we tested
SSD, HDDs, NVMe devices and discovered a wide range of behaviors and
performance profiles. As an example, Figure~\ref{fig:jewel-hdd-128b} shows the
write performance of 128 byte log entries using Jewel and a single HDD.
Performance is 10\(\times\) slower than its SSD counterpart in
Figure~\ref{fig:jewel_v_firefly_v_es} (top row, third column) but the behavior
and relative performance make this hardware configuration especially tricky.

The behavior of the 1:1 implementations shows throughput drops lasting for
minutes at a time -- this limits our focus to the N:1 implementations. The
performance of \texttt{(N:1, BS)} implementation is almost identical to
\texttt{(N:1, KV)} (within 1\% mean throughput). Regardless of the true
bottleneck, it is clear that choosing \texttt{(N:1, KV)} is the better choice
because of the resulting implementation should be less complex and there is
minimal performance degradation.
